---
title: "Merge, descripive and statistical analysis"
author: "Nelly"
date: "2024-03-15"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r}
library(readxl)
library(grabr)
library(gagglr)
library(tidyverse)
library(glamr)
library(glitr)
library(gophr)
library(systemfonts)
library(scales)
library(tidytext)
library(glue)
library(janitor)
library(ggtext)
library(patchwork)
library(knitr)
library(readr)
library(openxlsx)
library(stringr)
```

# set filepath
```{r}
# data_folder <- "Data"
# data_folder %>% return_latest()
ml_path <- "../Dataout/masterclientlist.rds"
ml_dfclean <- read_rds(ml_path)
```

```{r}
# data_folder <- "Data"
# data_folder %>% return_latest()
vl_path <- "../Dataout/vldata.rds"
vl_dfclean <- read_rds(vl_path)
```



```{r}
vl_dfclean %>% 
    group_by(merge_id) %>% 
    filter(n() > 1)
```



```{r}
dupl <- vl_dfclean[vl_dfclean %>% 
    select(merge_id, datetested) %>% 
    duplicated(), ] 
```

Remove duplicates

```{r}
#remove duplicates
vl_dfclean <- vl_dfclean %>% 
  group_by(merge_id, datetested) %>% 
  filter(n() == 1) %>% 
  ungroup()

#how many observations are there for each unique merge id (there should be 3)
vl_dfclean %>% 
  count(merge_id, name = "stages") %>% 
  count(stages) %>% 
  mutate(share = n /sum(n))

#remove where only 1 observation recorded
vl_dfclean <- vl_dfclean %>% 
  group_by(merge_id) %>% 
  filter(n() > 1) %>% 
  ungroup()
# check the CALHIV with more viral loads 
```

@ Qc to see if the outliers were data issues 
```{r}
# Count the number of observations for each merge_id
observations <- vl_dfclean %>% 
  count(merge_id, name = "stages")

# Filter to find merge_ids that occur exactly 7 or 8 times
observations_seven_eight <- observations %>% 
  filter(stages == 7 | stages == 8)
#
```



Limit each patient to only 3 observations (max) and identify each stage

```{r}
vl_dfclean %>% 
  count(merge_id) %>% 
  count(n)

### could think about introducing some logic to ensure visits are 6 months from the last to possibly deal with some of the extra observations

#arrange each merge id by date and take the 1st, 2nd, and last (renaming it 3rd for uniformity purposes and drop anything else in between)
vl_dfclean_lim <- vl_dfclean %>% 
  group_by(merge_id) %>% 
  arrange(merge_id, datetested) %>% 
  mutate(stage = row_number(),
         group = case_when(n() == 2 ~"only 2 observations",
                           n() == 3 ~ "3 observations",
                           TRUE ~ "3+ observations")
         ) %>% 
  filter(stage <= 2 | stage == max(stage)) %>% #this takes the last obs but could take the 3rd as final instead
  ungroup() %>%
  mutate(stage = ifelse(stage > 2, 3, stage)) 
#histogram 


```
# Distribution  of VL per CALHIV

```{r}

observations_per_id <- vl_dfclean %>%
  group_by(merge_id) %>%
  summarize(num_observations = n()) %>%
  ungroup()

# Define binwidth
binwidth_value <- 1

# Create histogram with bars and labels
gg <- ggplot(observations_per_id, aes(x = num_observations)) +
      geom_histogram(aes(y = ..count..), binwidth = binwidth_value, fill = hw_viking, color = "black") +
      geom_text(
        aes(label = scales::percent(..count.. / sum(..count..), accuracy = 0.1)),  # Rounds to one decimal place
        stat = "bin", 
        binwidth = binwidth_value, 
        vjust = -0.5, 
        color = "black"
      ) +
      labs(title = "Distribution of Number of Viral Load Tests per CALHIV",
           x = "Number of Viral Load Tests", 
           y = "CALHIV") +
      si_style_ygrid() +
      ylim(0, NA)  # Set the y-axis limits to automatically accommodate the highest bar

# Print the plot
print(gg)
```

```{r}

# Define binwidth
binwidth_value <- 1

# Create histogram with bars and labels, excluding 0% labels and with conditional fill colors
gg <- ggplot(observations_per_id, aes(x = num_observations)) +
      geom_histogram(aes(y = ..count.., fill = ifelse(num_observations == 2, "grey", hw_viking)), binwidth = binwidth_value, color = "black") +
      geom_text(
        aes(label = ifelse(..count.. / sum(..count..) == 0, "", scales::percent(..count.. / sum(..count..), accuracy = 0.1))), 
        stat = "bin", 
        binwidth = binwidth_value, 
        vjust = -0.5,
        color = "black"
      ) +
      scale_fill_identity() +  # Use the colors directly
      labs(title = "Distribution of Number of Viral Load Tests per CALHIV",
           subtitle = "Over 50% of the CALHIV had more than 3 VLtests done over the project period",
           x = "Number of Viral Load Tests", 
           y = "CALHIV") +
        si_style_ygrid() +
      ylim(0, NA)  # Set the y-axis limits to automatically accommodate the highest bar

# Print the plot
print(gg)
```



Reshape wide to have one row per patient

```{r}
vl_df_clean_wide <- vl_dfclean_lim %>% 
  pivot_wider(names_from = stage,
              names_prefix = "V",
              values_from = c(result, datetested, vlsup_cat)) %>% 
  mutate(missing_V3 = is.na(result_V3),
         result_V3 = ifelse(is.na(result_V3), result_V2, result_V3),
         datetested_V3 = case_when(is.na(datetested_V3) ~ datetested_V2, TRUE ~ datetested_V3),
         vlsup_cat_V3 = ifelse(is.na(vlsup_cat_V3), vlsup_cat_V2, vlsup_cat_V3),
         )
```




## Perform a full join on ml_dfclean and vl_dfclean  using a common key'merge_id'
```{r}

#df1 <- tidylog::full_join(ml_dfclean, vl_df_clean_wide, by = "merge_id")

```
# remove 343 clients observations without VL data 
```{r}
# df1_filtered <- df1 %>%
#   dplyr::filter(!is.na(vlsup_cat_V1) | !is.na(vlsup_cat_V3))
```

Recoding the non intervention group 
```{r}
# df1_filtered <- df1_filtered %>% 
#   mutate(intervention = ifelse(is.na(intervention), 0, 1))

```



```{r}
# Recode specified levels into a single category "Relatives"
ml_dfclean <- ml_dfclean %>%
  mutate(parenthood_status_combined = case_when(
    parenthood_status %in% c("Work place", "wife", "Husband", 
                             "Grand Parents", "Child headed", 
                             "Caregiver mentor", "OVC") ~ "Relatives",
    TRUE ~ as.character(parenthood_status)
  ))
```



# recode VL variables to factors
```{r}
vl_df_clean_wide <- vl_df_clean_wide %>%
  mutate(across(c(vlsup_cat_V1, vlsup_cat_V2,vlsup_cat_V3), as.factor))
```


# adding intervention variable to the Vl df

```{r}

vl_df_cleanfinal <- vl_df_clean_wide %>%
  left_join(ml_dfclean %>% 
              distinct(merge_id, .keep_all = TRUE) %>%
              select(merge_id, intervention), 
            by = "merge_id") %>%
  mutate(intervention = replace_na(intervention, "No"),
         intervention = case_when(intervention == "Yes" ~ 1, 
                                  TRUE ~ 0))

```





#descriptive statistics  
```{r}

# Adjusting the subset to include only those with "Yes" in the intervention column but also one observation per merge_1d
intervention_subset <- ml_dfclean %>% 
  distinct(merge_id, .keep_all = TRUE) %>% 
  #filter(intervention == 1,) %>% 
  # Set the factor levels for the 'age' variable
  mutate(age = factor(age, 
                      levels = c("0-4 Years", "5-9 Years", "10-14 Years", "15-19 Years")))

moody_blue <- "#8980CB"  # New color code for moody blue
genoa <- "#287C6F"      # New color code for genoa

# Create the age-sex pyramid plot directly from the intervention subset
intervention_subset %>% 
  count(age, sex) %>% 
  mutate(count = ifelse(sex == "Female", -n, n),
         fill_color = ifelse(sex == "Female", moody_blue, genoa)) %>% 
  ggplot(aes(count, age, fill = fill_color)) +
  geom_blank(aes(x = -count)) + # center pyramind
  geom_col() +
  geom_text(aes(x = .75 * count, label = n)) + 
  scale_fill_identity() +
  labs(x = NULL, y = NULL, title = "Age-Sex Composition of the Intervention Group",
       subtitle = glue("<span style='color:{moody_blue}'>Female</span> v <span style='color:{genoa}'>Male</span> (N=640)"),
       caption = " Note:Includes all particlipants who participated in any of the assessments,
                            Source:PCM 2024 data") +
   si_style_ygrid()+
 # si_style_nolines() +
  theme(axis.text.x = element_blank(),
        plot.subtitle = element_markdown())
       

```


 # DISTRIBUTION by caregiver
 
```{r}
# Convert the combined parenthood status to a factor
ml_dfclean$parenthood_status_combined <- factor(ml_dfclean$parenthood_status_combined)


#create bar plot
# count the number of each parenthood status
parenthood_counts <- ml_dfclean %>%
#filter(intervention == "1") %>%
  distinct(merge_id, .keep_all = TRUE) %>% 
count(parenthood_status_combined) %>%
mutate(parenthood_status_combined = fct_infreq(parenthood_status_combined)) %>%
mutate(percentage = n / sum(n) * 100) # Calculate percentages

# Plot
ggplot(parenthood_counts, aes(x = parenthood_status_combined, y = n, fill = parenthood_status_combined, label = scales::percent(percentage / 100))) +
 geom_bar(stat = "identity") +
  geom_text(vjust = -0.5, size = 3.5) + # Add percentage labels
  scale_fill_manual(values = c(grey10k, hw_sun_kissed)) +
  theme_minimal() +
 labs(x = "Parenthood Status", y = "# CALHIV", title = "Parenthood Status for Intervention Participants", subtitle = "67% of the participants had biological parents as caregivers") +
  si_style_ygrid()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x labels for readability
      legend.position = "none", # Hide legend
     panel.grid.major = element_blank(), # Remove major grid lines
     panel.grid.minor = element_blank(), # Remove minor grid lines
      plot.caption = element_text(hjust = 1)) + # Align caption to the right
 labs(caption = "Source: PCM 2024 data") # Add source caption

```
#categorizing vl data
```{r}
vl_df_cleanfinal <- vl_df_cleanfinal %>%
  # Convert factor levels to characters if they aren't already
  mutate(across(starts_with("vlsup_cat_V"), as.character)) %>%
  # Apply conditions based on 'group' and 'vlsup_cat' values
  mutate(vl_status = case_when(
    # If all instances are suppressed, then overall is suppressed
    group %in% c("3+ observations", "3 observations") & 
      vlsup_cat_V1 == "suppressed" & 
      vlsup_cat_V2 == "suppressed" & 
      vlsup_cat_V3 == "suppressed" ~ "suppressed",
      group %in% c("3+ observations", "3 observations") & 
      vlsup_cat_V1 == "unsuppressed" & 
      vlsup_cat_V2 == "suppressed" & 
      vlsup_cat_V3 == "suppressed" ~ "suppressed",
     group %in% c("3+ observations", "3 observations") & 
      vlsup_cat_V1 == "unsuppressed" & 
      vlsup_cat_V2 == "unsuppressed" & 
      vlsup_cat_V3 == "suppressed" ~ "suppressed",
    
    
    # If there is a mix of suppressed and unsuppressed, then overall is cycling
    group %in% c("3+ observations", "3 observations") & 
      vlsup_cat_V1 == "unsuppressed" & 
        vlsup_cat_V2 == "suppressed" & 
         vlsup_cat_V3 == "unsuppressed"~ "cycling",
        group %in% c("3+ observations", "3 observations") & 
      vlsup_cat_V1 == "suppressed" & 
        vlsup_cat_V2 == "suppressed" & 
         vlsup_cat_V3 == "unsuppressed"~ "cycling",
            group %in% c("3+ observations", "3 observations") & 
       (vlsup_cat_V1 == "suppressed" & 
          vlsup_cat_V2 == "unsuppressed" &
          vlsup_cat_V3 == "unsuppressed") ~ "cycling",
            group %in% c("3+ observations", "3 observations") & 
       (vlsup_cat_V1 == "suppressed" & 
          vlsup_cat_V2 == "unsuppressed" &
          vlsup_cat_V3 == "suppressed") ~ "cycling",
    
    # If all instances are unsuppressed, then overall is Never suppressed
    group %in% c("3+ observations", "3 observations") & 
      vlsup_cat_V1 == "unsuppressed" & 
      vlsup_cat_V2 == "unsuppressed" & 
      vlsup_cat_V3 == "unsuppressed" ~ "Never suppressed",
    
    # For the group "only 2 observations", apply similar logic for two visits
    group == "only 2 observations" & 
      vlsup_cat_V1 == "suppressed" & 
      vlsup_cat_V2 == "suppressed" ~ "suppressed",
        group == "only 2 observations" & 
      vlsup_cat_V1 == "unsuppressed" & 
      vlsup_cat_V2 == "suppressed" ~ "suppressed",
    group == "only 2 observations" & 
      vlsup_cat_V1 == "unsuppressed" & 
      vlsup_cat_V2 == "unsuppressed" ~ "Never suppressed",
    
    group == "only 2 observations" & 
      #(vlsup_cat_V1 == "unsuppressed" & vlsup_cat_V2 == "suppressed") | 
      (vlsup_cat_V1 == "suppressed" & vlsup_cat_V2 == "unsuppressed") ~ "cycling",
    
    # Default case if none of the above conditions are met
    TRUE ~ NA_character_  
  ))


```




```{r}
vl_df_cleanfinal %>%
  group_by(vl_status,intervention) %>%
  summarise(
    Count = n())

```



# exclude  all CALHIV in control group that had consistently been suppressed 

```{r}
vl_df_cleanfinal2 <- vl_df_cleanfinal %>%
  filter(intervention == 1 | (intervention == 0 & vl_status != "suppressed"))

```




#Adding age from the ml dataframe 
```{r}
first_age <- ml_dfclean %>%
  group_by(merge_id) %>%
  summarise(age = first(age))

# Merge the age column from first_age to vl_df_cleanfinal2 based on merge_id
vl_df_cleanfinal2 <- vl_df_cleanfinal2 %>%
  left_join(first_age, by = "merge_id")

```


```{r}
# Define your color
hw_lavender_haze <- "#E6E6FA" #control group 

long_data <- vl_df_cleanfinal2 %>%
  filter(missing_V3==FALSE) %>% 
  select(merge_id, intervention, vlsup_cat_V1, vlsup_cat_V2, vlsup_cat_V3) %>%
  pivot_longer(cols = c(vlsup_cat_V1, vlsup_cat_V2, vlsup_cat_V3), names_to = "time", values_to = "viral_load") %>%
  mutate(
    time = fct_recode(time, "Start" = "vlsup_cat_V1", "Midpoint" = "vlsup_cat_V2", "End" = "vlsup_cat_V3"),
    intervention = factor(intervention, labels = c("Control", "Intervention"))
  ) %>%
  filter(!is.na(viral_load))  # Exclude NA values

# Summarize the data to get counts
count_data <- long_data %>%
  group_by(time, intervention, viral_load) %>%
  summarise(count = n(), .groups = 'drop')

#Create the plot with specified facet layout and add count labels
ggplot(count_data, aes(x = time, y = count, fill = intervention)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count), position = position_dodge(width = 0.9), vjust = 1) +  # Add count labels
  facet_wrap(~ viral_load, scales = "free_y", nrow = 2) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.title = element_blank(),
          plot.caption = element_text(hjust = 1)) + # Align caption to the right
  labs(
    x = "",
    y = "# CALHIV",
    title = "Viral load by Time Point and Treatment Group",
    subtitle = "Viral Suppression consitently increased for the participants with 3 or more viral load tests in the group",  # Subtitle added here
    caption = " Note:Contol cohort was drawn from CALHIV who did not enrol in the intervetion at the site,
                            Source: PCM 2024 data"
  ) +
   si_style_ygrid()+
  scale_fill_manual(values = c("Control" = hw_lavender_haze, "Intervention" = "#9e94e0"), labels = c("Control", "Intervention")) +
  scale_x_discrete(labels = c("Start", "Midpoint", "End"))


```

#attempt to create a cohort heatmap 

```{r}
# Filter for intervention = 1 and missing_V3 is FALSE
filtered_df <- vl_df_cleanfinal2 %>%
  filter(intervention == 1, missing_V3 == FALSE)

# Summarize the total and suppressed counts for "vlsup_cat_V1"  "vlsup_cat_V2" "vlsup_cat_V
summary_table <- filtered_df %>%
  select(contains("vlsup_cat_V")) %>%  # Select only the relevant columns
  gather(key = "time_point", value = "suppression_status") %>%
  group_by(time_point, suppression_status) %>%
  summarise(count = n()) %>%
  spread(key = suppression_status, value = count) %>%
  mutate(Total = `suppressed` + `unsuppressed`)  # Add a column to calculate total counts
# Assuming your dataframe is named 'filtered_df'
summary_table <- filtered_df %>%
  select(contains("vlsup_cat_V")) %>%  # Select only the relevant columns
  gather(key = "time_point", value = "suppression_status") %>%
  group_by(time_point, suppression_status) %>%
  summarise(count = n()) %>%
  spread(key = suppression_status, value = count, fill = 0) %>%  # Fill NAs with 0
  mutate(Total = `suppressed` + `unsuppressed`,  # Calculate total counts
         Percentage_Suppressed = (`suppressed` / Total) * 100)  # Calculate percentage suppressed

# Convert to a format suitable for ggplot stacked bar chart
summary_table_long <- summary_table %>%
  mutate(unsuppressed = Total - suppressed) %>%
  mutate(Percentage_Suppressed = (suppressed / Total) * 100) %>%
  select(time_point, suppressed, unsuppressed, Percentage_Suppressed) %>%
  gather(key = "status", value = "count", -time_point, -Percentage_Suppressed) %>%
  group_by(time_point, status) %>%
  mutate(label = if_else(status == "suppressed", as.character(Percentage_Suppressed), ""))
# # Adjust the summary_table_long dataframe
summary_table_long <- summary_table_long %>%
  mutate(label = if_else(status == "suppressed", 
                         sprintf("%.0f%%", Percentage_Suppressed), ""))  # Round to two decimal places

# Reorder factor levels to have 'suppressed' at the bottom
summary_table_long$status <- factor(summary_table_long$status, levels = c("unsuppressed", "suppressed"))
summary_table_long$time_point <- factor(summary_table_long$time_point, 
                                        levels = c("vlsup_cat_V1", "vlsup_cat_V2", "vlsup_cat_V3"),
                                        labels = c("Start", "Mid", "End"))

# Plot the stacked bar chart without a legend and with renamed time points
stacked_bar <- ggplot(summary_table_long, aes(x = time_point, y = count, fill = status)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("unsuppressed" = "#E6E7E8", "suppressed" = "#004964")) +
    si_style_ygrid()+
 # theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Viral Suppression by Timepoint",
    subtitle = "Viral Suppression Increased Gradually Each Stage",
    caption = "Note:The data includes only CALHIV with 3+ vl tests
               Source:PCM data",
    fill = "Status",
    y = "# CALHIV",
    x = "",
     plot.title = element_markdown()
  
  )

# Print the plot
stacked_bar

```

# creating a panel chart to visualize age and viral suppression 

```{r}
# Filter out observations where age is NA
# Filter out observations where age is NA and set the factor levels for age
vl_df_cleanfinal2_filtered <- vl_df_cleanfinal2 %>%
  filter(!is.na(age)) %>%
  mutate(age = factor(age, levels = c("0-4 Years", "5-9 Years", "10-14 Years", "15-19 Years")))

```




```{r}

# First, summarize the counts of suppression categories within each age group and time point
age_timepoint_summary <- vl_df_cleanfinal2_filtered %>%
  gather(key = "time_point", value = "suppression_status", starts_with("vlsup_cat_V")) %>%
  group_by(age, time_point, suppression_status) %>%
  summarise(count = n(), .groups = "drop") %>%
  # Calculate the percentage of each suppression status within each age group and time point
  group_by(age, time_point) %>%
  mutate(total = sum(count),
         percent = (count / total) * 100) %>%
  # Spread the data for plotting
  ungroup() %>%
  select(-total, -count) %>%
  spread(key = suppression_status, value = percent)
age_timepoint_summary <- age_timepoint_summary %>%
  mutate(time_point = factor(time_point, levels = c("vlsup_cat_V1", "vlsup_cat_V2", "vlsup_cat_V3"),
                             labels = c("Start", "Midpoint", "End")))

# Next, create the diverging stacked bar chart
ggplot(age_timepoint_summary, aes(x = age, y = unsuppressed, fill = "unsuppressed")) +
  geom_bar(stat = "identity") +
  geom_bar(aes(y = -suppressed, fill = "suppressed"), stat = "identity") +
  geom_text (color= "white",aes( y = -suppressed/2,label= scales::percent(abs(suppressed/100),accuracy = 1))) +
    facet_wrap(~time_point, scales = "free_x") +
  coord_flip() +
  scale_fill_manual(values = c("suppressed" = "#004964", "unsuppressed" = "#D5D7D8")) +
  scale_y_continuous(labels = abs) +  # To display positive labels on both sides
  labs(
    x = "",
    y = "Percentage",
    fill = "Suppression Status",
    title = "Viral <span style='color:#004964;font-weight:1000'>Suppression</span> by Age and Stage",
    subtitle =  "Viral suppression consistently increased at each stage",
    caption = " Note:Includes only participants in the intervention
                Source:PCM Uganda data",) +
 si_style_ygrid()+
  #theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_markdown())
        # si_style_nolines()

```

```{r}
ggplot(age_timepoint_summary, aes(x = age, y = suppressed, fill = "suppressed")) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(suppressed), "%")), 
            position = position_stack(vjust = 0.5), 
            color = "white", size = 3) +  # Add rounded percentage labels
  coord_flip() +
  scale_fill_manual(values = c("suppressed" = "#004964")) +
  labs(
    x = "",
    y = "Percentage of Suppressed",
    fill = "Suppression Status",
    title = "Viral Suppression by Age and Timepoint",
      subtitle = "All age bands showed improvement over the mid and end points",
    caption = "Source: PCM Uganda data"
  ) +
  si_style_ygrid()+
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    panel.grid.major.x = element_blank(),  # Remove x-axis gridlines
    axis.title.y = element_text(margin = margin(r = 20), angle = 0, vjust = 0.5),  # Adjust y-axis label position
   # axis.line.y = element_line(color = "black"),  # Add y-axis line
    axis.ticks.y = element_blank(),  # Remove y-axis ticks
    #axis.text.y = element_blank()  # Remove y-axis text
  ) +
  facet_wrap(~time_point, scales = "free_y")
```





```{r}


# Filter the data to include only 'suppressed' viral_load status
suppressed_data <- long_data %>%
  filter(viral_load == "suppressed") %>%
  count(intervention, time) # count occurrences of 'suppressed' for each intervention and time

# Check the structure to ensure it's correct before plotting
print(suppressed_data)

# # Create the heatmap for 'suppressed' status
# ggplot(suppressed_data, aes(x = time, y = intervention, fill = n)) +
#   geom_tile(color = "white") +  # Use white borders for the tiles
#   scale_fill_gradient(low = "lavender", high = "blue") +  # Use a gradient from light to dark for the counts
#   labs(
#     x = "Time Point",
#     y = "Intervention Group",
#     fill = "Count",
#     title = "Change in Viral Load Status Over Time",
#     subtitle = "Count of Suppressed Status"
#   ) +
#   theme_minimal() +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     axis.text.y = element_text(angle = 45, hjust = 1),
#     axis.title = element_text(size = 12)
#   )


```
# structuring the data for sankey visual 


```{r}
library(ggalluvial)

```





#analysis for CALHIV with 2 blood Viral loads 

```{r}
long_data <- vl_df_cleanfinal2 %>%
  filter(missing_V3==TRUE) %>% 
  select(merge_id, intervention, vlsup_cat_V1, vlsup_cat_V3) %>%
  pivot_longer(cols = c(vlsup_cat_V1, vlsup_cat_V3), names_to = "time", values_to = "viral_load") %>%
  mutate(
    time = fct_recode(time, "Start" = "vlsup_cat_V1",  "End" = "vlsup_cat_V3"),
    intervention = factor(intervention, labels = c("Control", "Intervention"))
  ) %>%
  filter(!is.na(viral_load))  # Exclude NA values

# Summarize the data to get counts
count_data <- long_data %>%
  group_by(time, intervention, viral_load) %>%
  summarise(count = n(), .groups = 'drop')

#Create the plot with specified facet layout and add count labels
# Define your color
hw_lavender_haze <- "#E6E6FA"  # Example lavender color, adjust the code to match your desired shade


ggplot(count_data, aes(x = time, y = count, fill = intervention)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = count), position = position_dodge(width = 0.9), vjust = 1) +  # Add count labels
  facet_wrap(~ viral_load, scales = "free_y", nrow = 1) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.title = element_blank()
  ) +
  labs(
    x = "Time",
    y = "CALHIV",
    title = "Viral load by Time Point and and Treatment Group",
    subtitle = "Viral Suppression increase over time for the participants with 2 viral load tests" , # Subtitle added here,
    caption = "Source:PCM Uganda data"
  ) +
  scale_fill_manual(values = c("Control" = hw_lavender_haze, "Intervention" = "#00BFC4"), labels = c("Control", "Intervention")) +
  scale_x_discrete(labels = c("Start",  "End"))

```

#creating final anaysis dataframe 
```{r}
# df_stat<- df_stat <- df1_filtered %>%
#   select(
#     merge_id,
#     age,
#     sex,
#     group,
#     recommend_services,
#     healthy_symptom_free_frequency,
#     problem_experience_last_2_weeks,
#     support_from_family_friends,
#     hiv_discrimination_past_month,
#     factors_impacted_score_most,
#     hiv_treatment_expenses_loan_sold,
#     preferred_hiv_services_method,
#     attach_to_chw,
#     parenthood_status_combined,
#     intervention,
#     vlsup_cat_V1,
#     vlsup_cat_V2,
#     vlsup_cat_V3,
#   )
```




```{r}

# Your existing code to create the summary table
summary_data <- vl_df_cleanfinal2 %>%
  group_by(vl_status, intervention) %>%
  summarise(Count = n(), .groups = 'drop')  # Ensure to drop grouping

# # Creating a bar plot
# ggplot(summary_data, aes(x = intervention, y = Count, fill = vl_status)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   theme_minimal() +
#   labs(x = "Intervention", y = "Count", title = "Count of VL Outcomes by Intervention") +
#   scale_fill_brewer(palette = "Set1") +  # Optional: Use a color palette that works well for categorical data
#   theme(plot.title = element_text(hjust = 0.5))  # Center the plot title

```
# creating a treemap 
```{r}
library(treemapify)


# ggplot(summary_data, aes(area = Count, fill = vl_status, label = paste(vl_status, Count))) +
#   geom_treemap() +
#   geom_treemap_text(fontface = "italic", colour = "white", place = "centre", grow = TRUE) +
#   facet_wrap(~ intervention, scales = "free") +
#   labs(title = "Treemap of VL Outcome Counts by Intervention") +
#   theme_minimal() +
#   theme(plot.title = element_text(hjust = 0.5))  # Center the plot title
```

```{r}
summary_data$intervention <- factor(summary_data$intervention, levels = c(0, 1), labels = c("Control", "Intervention"))

#summary_data$intervention <- factor(summary_data$intervention, levels = c(0, 1), labels = c("Control", "Intervention"))

# First, calculate the total counts for each intervention group to use for percentage calculation
summary_data <- summary_data %>%
  group_by(intervention) %>%
  mutate(Total = sum(Count)) %>%
  ungroup() %>%
  mutate(Percentage = Count / Total * 100)  # Calculate percentages


```



```{r}

# Now create the heatmap with percentage labels
ggplot(summary_data, aes(x = vl_status, y = as.character(intervention), fill = Count)) +
  geom_tile() +  # Use geom_tile() for heatmap squares
  scale_fill_gradient(low = "white", high = "#7f001c") +  # Color gradient for the count
  geom_text(aes(label = paste0(round(Percentage, 0), "%")), color = "black", size = 3.5) +  # Add percentage labels
  labs(x = "VL_status category", 
       y = "", 
       title = "Comparison of Viral Load Outcomes by Group",
       subtitle ="69% of the participants in the intervention group were cycling compared to 79% in the control",
       caption = "Note:Data includes  Intervention and control
                    Data excludes CALHIV who were suppressed though out the project period
                  Source:PCM Uganda data 2024",
       fill = "Count") +
geom_hline(yintercept = 1.5,linewidth = 2, color = "white")+
   si_style_ygrid()+
  #theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0),  # Center the plot title
    legend.position = ""  # Move the count legend to the bottom
  )

```

# Alluvial 
```{r}
alluv_df<- vl_df_cleanfinal2_filtered %>% 
  group_by(vlsup_cat_V1, 
           vlsup_cat_V2,vlsup_cat_V3,vl_status,) %>% 
  summarise(n=n())

```

```{r}

alluv_df <- vl_df_cleanfinal2_filtered %>%
  filter(intervention == 1) %>%
  group_by(vlsup_cat_V1, vlsup_cat_V2, vlsup_cat_V3, vl_status) %>%
  summarise(n = n(), .groups = 'drop')  # Adding .groups = 'drop' to prevent the creation of an extra grouping layer

```






```{r}
ggplot(data = alluv_df,
       aes(axis1 = vlsup_cat_V1, axis2 = vlsup_cat_V2, axis3 = vlsup_cat_V3, axis4 = vl_status, y =  n)) +
  geom_alluvium(#data = alluv_df %>% filter( vl_status == "cycling"), 
                aes(fill = vl_status)) +
  geom_stratum(aes(fill = vl_status)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)))+
  scale_fill_manual(values = c("suppressed" = grey20k, "cycling" = hw_orchid_bloom, "Never suppressed" = hw_midnight_blue)) +
    scale_x_discrete(labels= c("Start","Midpoint","End", "VL Outcome")) +
si_style_nolines()+
theme(legend.position = "none")+
labs(title = "VL Suppression Outcome for CALHIV with 3 VL Loads",
     subtitle = "25% of the participants never suppressed",
      # subtitle = "Cycling Patients at 16% of the Patients in the Intervention Group",
        y = "# of CALHIV",
       caption = "Note:Includes only intervention cohort
                  Source:PCM Data Uganda")

 # scale_fill_viridis_d()
 # scale_fill_brewer(type = "div", palette = "Set1") 
 # theme_minimal()

```



```{r}
ggplot(data = alluv_df, aes(axis1 = vlsup_cat_V1, axis2 = vlsup_cat_V2, axis3 = vlsup_cat_V3, axis4 = vl_status, y = n)) +
  geom_alluvium(aes(fill = vl_status)) +
  geom_stratum(aes(fill = vl_status)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_fill_manual(values = c("suppressed" = grey20k, "cycling" = hw_orchid_bloom, "Never suppressed" = grey20k)) +
  scale_x_discrete(labels= c("Start","Midpoint","End", "VL Outcome")) +
  si_style_nolines() +
  theme(legend.position = "none") +
  labs(title = "VL Suppression Outcome for Patients with 3 VL Loads",
       subtitle = "Cycling Patients at 16% of the Patients in the Intervention Group",
       caption = "Note:Includes only intervention cohort
                  Source:PCM Data Uganda")


```
```{r}

# library(dplyr)
# 
# # Create the stratum_label column
# alluv_df <- alluv_df %>%
#   mutate(stratum_label = vl_status)
# 
# # Now plot with ggplot2
# ggplot(data = alluv_df, aes(axis1 = vlsup_cat_V1, axis2 = vlsup_cat_V2, axis3 = vlsup_cat_V3, axis4 = vl_status, y = n)) +
#   geom_alluvium(aes(fill = vl_status)) +
#   geom_stratum(aes(fill = vl_status)) +
#   geom_text(stat = "stratum", aes(label = after_stat(stratum)), position = position_stack(vjust = 0.5)) +
#   scale_fill_manual(values = c("suppressed" = "grey", "cycling" = "blue", "Never suppressed" = "grey")) +
#   scale_x_discrete(labels = c("Start", "Midpoint", "End", "VL Outcome")) +
#   theme(legend.position = "none") +
#   labs(title = "VL Suppression Outcome for Patients with 3 VL Loads",
#        subtitle = "Cycling Patients at 16% of the Patients in the Intervention Group",
#        caption = "PCM Data Uganda")


```


```{r}


# Define colors for the levels of 'vl_status'
colors <- c("suppressed" = "red", "cycling" = "blue", "unsuppressed" = "grey", "Never suppressed" = "black")

geom_alluvium(aes(fill = vl_status, alpha = 0.5))  # sets alpha to 50%


# Assuming 'alluv_df' is your dataframe and it's properly structured with the columns:
# 'vlsup_cat_V1', 'vlsup_cat_V2', 'vlsup_cat_V3', 'vl_status', and 'n'

ggplot(data = alluv_df,
       aes(axis1 = vlsup_cat_V1, axis2 = vlsup_cat_V2, axis3 = vlsup_cat_V3, axis4 = vl_status, y = n)) +
  geom_alluvium(aes(fill = vl_status), alpha = 0.5) + # Set alpha directly here
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()

```

```{r}

# Set the path to the folder where you want to save the CSV
output_folder <- "C:/Users/nmaina/Documents/R.Projects/UgandaMatrixMix/Dataout"

# Define the full path to the CSV file within the output folder
output_file_path <- file.path(output_folder, "Alluv_data.csv")

# Write the vl_df1 DataFrame to the CSV file
write.csv(alluv_df, output_file_path, row.names = FALSE)

```

