---
title: "PCM_ uganda_2024"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Load packages 
```{r}
library(readxl)
library(grabr)
library(gagglr)
library(tidyverse)
library(glamr)
library(glitr)
library(gophr)
library(extrafont)
library(scales)
library(tidytext)
library(glue)
library(janitor)
library(ggtext)
library(extrafont)
library(patchwork)
library(knitr)
library(readr)
library(openxlsx)
```

```{r}
# setwd("C:/Users/nmaina/Documents/R.Projects/PCM_ uganda_2024")
```

# set filepath
```{r}
# data_folder <- "Data"
# data_folder %>% return_latest()
file_path <- "Data/PCM _Acholi 26 Feb.xlsx"
```


Read data 
```{r}
df <- read_excel(file_path,
                 "Other Parameters")
```


clean master clientlist

rename variables 
```{r}
# new variable names
names(df) <- c("interview_date", "district", "facility", "client_art", "merge_id", "assessment_type", "setting", "age", 
               "sex", "art_duration", "appt_missed_last_12_months", "recommend_services", "q1_impact",
               "healthy_symptom_free_frequency", "symptoms_impacted_q1", "experienced_symptoms_last_2_weeks", 
               "problem_experience_last_2_weeks", "support_from_family_friends", "hiv_discrimination_past_month",
               "factors_impacted_score_most", "hiv_treatment_expenses_loan_sold", "specific_expenses_for_treatment",
               "art_dispensed_last_contact", "preferred_art_dispensed", "unsuppressed_calhiv_dsd_approach",
               "preferred_hiv_services_method", "preferred_dsd_choice", "attach_to_chw", "parenthood_status")

```

#recode Interview questions  
q1_impact 
```{r}
df <- df %>% 
  mutate(q1_impact =  recode(q1_impact,
                             "Not interested" = "No data",
                             "Minor" = "No data",
                             "Drugs out of stock" = "Other",
                             "No youths corner" = "Other",
                             "its hard to swallow ART Drugs daily." = "Other",
                             "Counselling given from facility has helped alot" = "Access to useful information",
                             "Adherence counselling" = "Access to useful information",
                             "She is a transfer out to TASO Gulu" =  "Other",
                             "Self stigma" = "Other",
                             "None of the above" = "Other",
                             .default = df$q1_impact
  ))

```

symptoms 
```{r}
df <- df %>% 
  mutate(factors_impacted_score_most_aggregated = factors_impacted_score_most %>% 
           fct_lump(n = 4) %>% 
           as.character) 
```




```{r}

df <- df %>%
  mutate(symptoms_2w_family = str_detect(experienced_symptoms_last_2_weeks, "Belief and behavior of your family toward you"),
         symptoms_2w_community = str_detect(experienced_symptoms_last_2_weeks, "Belief and behavior of the community toward you"),
         symptoms_2w_staff = str_detect(experienced_symptoms_last_2_weeks, "Attitude and behavior of clinic staff toward you"),
         )


```


```{r}
# Define a vector of variable names that need type conversion
factor_columns <- c("district", "facility", "assessment_type", "setting", "age", "sex", "art_duration",
                    "appt_missed_last_12_months", "recommend_services", "q1_impact", "healthy_symptom_free_frequency", "problem_experience_last_2_weeks",
                    "support_from_family_friends", "hiv_discrimination_past_month", "factors_impacted_score_most",
                    "hiv_treatment_expenses_loan_sold", "specific_expenses_for_treatment", "art_dispensed_last_contact",
                    "preferred_art_dispensed", "unsuppressed_calhiv_dsd_approach", "preferred_hiv_services_method",
                    "preferred_dsd_choice", "attach_to_chw", "parenthood_status")

df <- df %>%
  mutate(client_art = as.character(client_art),
         merge_id = as.character(merge_id),
         interview_date = as_date(interview_date))

# Check the structure of the modified data frame


df <- df %>%
  mutate_at(vars(all_of(factor_columns)), as.factor)

# Check the structure of the modified data frame
str(df)

```
Remove  variables not needed for analysis 

```{r}
df<- df %>%
  select(-setting,-district,-facility, -art_duration, -appt_missed_last_12_months, -client_art)
```


# looking for duplicates deleted 4 duplicate observations
```{r}
# Remove duplicate rows, keeping only the first occurrence
df_clean <- distinct(df)

```

# looking for duplicates based on merge_id, assessment_type, interview_date
```{r}
# Identify duplicates based on merge_id, assessment_type, and interview_date
duplicates <- df_clean[duplicated(df_clean[c("merge_id", "assessment_type", "interview_date")]) | 
                  duplicated(df_clean[c("merge_id", "assessment_type", "interview_date")], fromLast = TRUE), ]
```



Removing second occurrence of duplicate
```{r}
df_clean2<- df_clean %>%
  group_by(merge_id, assessment_type, interview_date) %>%  # Group by the specified columns
  filter(row_number() == 1) %>%  # Keep only the first occurrence
  ungroup()  # Remove the grouping
```

13 duplicates were identified, 
```{r}
# Selecting non-duplicate rows based on 'merge_id' and 'assessment_type'
df_clean3 <- df_clean2[!duplicated(df_clean2[c("merge_id", "assessment_type")]) & 
                                    !duplicated(df_clean2[c("merge_id", "assessment_type")], fromLast = TRUE), ]

# df_clean3 will contain data with the second and subsequent duplicates removed

```

#The resulting df4 will contain 1442 all observations from df_clean3 for each merge_id that are associated with more than one assessment_type. these will be the master c

```{r}
df4 <- df_clean3 %>%
  group_by(merge_id) %>%
  filter(n_distinct(assessment_type) > 1) %>%
  ungroup()
```


<!-- #creates a new df  single_assessment. The df will have observations by merge_id with only one assessment type  -->
<!-- ```{r} -->
<!-- single_assessment <- df_clean3 %>% -->
<!--   group_by(merge_id) %>% -->
<!--   filter(n_distinct(assessment_type) == 1) %>% -->
<!--   ungroup() -->

<!-- # single_assessment will contain 425 observations by merge_id with only one assessment_type -->
<!-- ``` -->



#This code adds a new column intervention to df4 and assigns the value "Yes" to this column for every row in the DataFrame. this is to identify all the observations in the intervention group 
```{r}
df4 <- df4 %>%
  mutate(intervention = "Yes")
```

Checking whether all observations have merge_ids

```{r}
df4_summary <- df4 %>%
  summarise(total_rows = n(),
            non_missing_assessment_type = sum(!is.na(assessment_type)))
```


# Write the df4 DataFrame to the CSV file
```{r}
# Set the path to the folder where you want to save the CSV
output_folder <- "C:/Users/nmaina/Documents/R.Projects/PCM_ uganda_2024/dataout"

# Define the full path to the CSV file within the output folder
output_file_path <- file.path(output_folder, "masterclientlist.csv")

# Write the vl_df1 DataFrame to the CSV file
write.csv(df4, output_file_path, row.names = FALSE)

```

